Delivering Reliable Software in the European Scientific Arena: Evolution of Methodologies and New Trend Analysis

### 2017-12-01 paper-ieee-trans.tex
added changes

\author{E. Ronchieri,
        P. Orviz Fernandez,
        M. David,
		D. C. Duma,
        J. Gomes,
        D. Salomoni
\thanks{D. C. Duma, E. Ronchieri, and D. Salomoni are with INFN - CNAF, Bologna, Italy.}
\thanks{P. Orviz Fernandez, is with CSIC, Santander, Spain, (e-mail: orviz@ifca.unican.es)}
\thanks{M. David, and J. Gomes are with LIP, Lisbon, Portugal}%
}

@@

\IEEEPARstart{I}{n} the European scientific research arena, the analysis of
some of the last decade software development projects depicts a continuous
increase of the prominence and robustness of testing and validation procedures.
The observed trends are aligned with the evolution of the software engineering
techniques (e.g. rise of DevOps practices \cite{zhu}) over the last years. This is tied to
the parallel advancements in the ICT field (with virtualization and container
technologies) and the emergence of automation and event-response tools.

>>>>

\author{E. Ronchieri,
        P. Orviz Fernandez,
        M. David,
		D. C. Duma,
        J. Gomes,
        D. Salomoni
\thanks{D. C. Duma, E. Ronchieri, and D. Salomoni are with INFN - CNAF, Bologna, Italy.}
\thanks{P. Orviz Fernandez, is with CSIC, Santander, Spain, (e-mail: orviz@ifca.unican.es)}
\thanks{M. David, and J. Gomes are with LIP, Lisbon, Portugal.}%
}

@@

\IEEEPARstart{A}{n}alyzing the European Commission-funded software projects in the last decade we have observed a continuous
increase of the prominence and robustness of testing and validation procedures. The first of many reasons is related to the evolution of the software engineering
techniques (e.g. rise of DevOps practices \cite{zhu}) together with the parallel advancements in the ICT field with virtualization and container
technologies. Another one is the emergence of automation and event-response tools.

Software developers involved in these projects have learnt over time how to make their product more reliable in order to satisfy requirements of different communities, such as physics and bio-informatics.
